---
title: "acf-pcf"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{acf-pcf}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup}
library(tsibble)
library(dplyr)
library(fable)
library(tsibbledata)
library(ggplot2)
library(lubridate)
library(feasts)
library(tidyr)
```

## chapter 9 refresher

### stationary
```{r}
google_2015 <- gafa_stock %>%
  filter(Symbol == "GOOG") %>%
  mutate(day = row_number()) %>%
  update_tsibble(index = day, regular = TRUE) %>%
  filter(year(Date) == 2015)
```

```{r}
google_2015 %>%
  gg_tsdisplay()
```


ARIMA needs to work on stationary data

*stationary*
- no trends or seasonality
- cyclic behavior okay
- white noise is an example, but so is lynx



### differences `d` parameter
d parameter: number of differences until data is stationary

Test if stationary
```{r}
google_2015 %>%
  mutate(diff_close = difference(Close)) %>%
  features(diff_close, ljung_box, lag = 10)
```

Find number of differences:
```{r}
google_2015 %>%
  features(Close, unitroot_ndiffs)
```


### autogressive `p` parameter

Autoregressive parameter `p`


### moving average `q` parameter

moving average `q` parameter





### ARIMA

`pdq`

ARIMA(0, 0, 0)     ~ white noise
ARIMA(0, 1, 0) - 1 ~ random walk (one difference)
ARIMA(0, 1, 0) + 1 ~ random walk with drift (one difference)


#### example

```{r}
us_change <- readr::read_csv("https://otexts.com/fpp3/extrafiles/us_change.csv") %>%
  mutate(Time = yearquarter(Time)) %>%
  as_tsibble(index = Time)
```

```{r}
us_change %>% 
  gg_tsdisplay(Consumption)
```

```{r}
fit <- us_change %>%
  model(ARIMA(Consumption ~ PDQ(0,0,0)))

report(fit)
```



### Using ACF and PCF with ARIMA

We can use ACF and PACF (of the difference `d` data)
to determine order of `p` and `q`.

ARIMA(p, d, 0)
`p` ->
  * spike on lag `p` of PACF, none past `p`
  * ACF sinusodial or decaying
  
  Seasonal: `p` is the number of spikes at at the seasonal interval on PACF
  
  > `p` is found on using the PACF partial autocorrelation because the effects
    due to short-term intervals are removed. 
    This finds the "true" autocorrelation 

ARIMA(0, d, q)
`q` ->
  * spike on lag `q` of ACF, none past `q`
  * PACF sinusodial or decaying
  
  Seasonal: `q` is the number of spikes at at the seasonal interval on ACF
  
  > `q` uses ACF, because if some sequential lags are autocorrelated,
    then the moving average will be able to use those successive observations
    to build a better moving average model.
  
```{r}
us_change %>% ACF(Consumption) %>% autoplot()
```

```{r}
us_change %>% PACF(Consumption) %>% autoplot()
```


```{r}
eu_retail <- as_tsibble(fpp2::euretail)
eu_retail %>% autoplot(value) + ylab("Retail index") + xlab("Year")
```


```{r}
eu_retail %>% gg_tsdisplay(value %>% difference(4) %>% difference(),
  plot_type='partial')
```

Try suggested model:

```{r}
fit <- eu_retail %>%
  model(arima = ARIMA(value ~ pdq(0,1,1) + PDQ(0,1,1)))
fit %>% gg_tsresiduals()
```

We see a spike on ACF lag 2 and 3. This suggests should should increase the 
order of the moving to 3 too account for those terms.

```{r}
fit <- eu_retail %>%
  model(ARIMA(value ~ pdq(0,1,3) + PDQ(0,1,1)))
fit %>% gg_tsresiduals()
```

Which results in a better fit, without autocorrelations.

#### Example 2

```{r}
h02 <- tsibbledata::PBS %>%
  filter(ATC2 == "H02") %>%
  summarise(Cost = sum(Cost)/1e6)
h02 %>%
  mutate(log(Cost)) %>%
  gather() %>%
  ggplot(aes(x = Month, y = value)) +
  geom_line() +
  facet_grid(key ~ ., scales = "free_y") +
  xlab("Year") + ylab("") +
  ggtitle("Cortecosteroid drug scripts (H02)")
```

We choose `log(Cost)` because there is less seasonal variance. 

```{r}
h02 %>% 
  gg_tsdisplay(difference(log(Cost), 12), plot_type='partial', lag_max = 24)
```

The autocorrelation plots strongly suggest an `ARIMA(3,0,0)(2,1,0)`.
- 1 seasonal difference
- 0 zero term difference



## Autocorrelation 2.8

What is ACF?

> autocorrelation measures the linear relationship between lagged values 
of a time series.


Trends
> When data have a trend, the autocorrelations for small lags tend to be large 
and positive because observations nearby in time are also nearby in size.



```{r}
recent_production <- tsibbledata::aus_production %>%
  filter(year(Quarter) >= 1992)

recent_production %>%
  gg_tsdisplay()
```

```{r}
recent_production %>%
  ACF(Beer)
```

```{r}
recent_production %>%
  ACF(Beer) %>%
  autoplot()
```

ACF allows use to see season correlations.
- a lag of 4 because that is the seasonal period
- negative lag of two, 
  because that is when the it turns from a trough to a peak.


```{r}
a10 <- tsibbledata::PBS %>%
  filter(ATC2=="A10") %>%
  select(Month, Concession, Type, Cost) %>%
  summarise(TotalC = sum(Cost)) %>%
  mutate(Cost = TotalC/1e6)
```

```{r}
a10 %>%
  gg_tsdisplay()
```

In a10, we see a increase trend, and also seasonality in Janurary.

```{r}
a10 %>%
  ACF(Cost, lag_max = 48) %>%
  autoplot()
```

We see a decrease correlation as the series progresses, 
because there is an increasing trend.

However, we do see peaks at 12, 24 (intervals of the seasonlity).





## next

