---
title: "fable-chapter-3"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{fable-chapter-3}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r}
library(tsibble)
library(dplyr)
library(fable)
library(tsibbledata)
library(ggplot2)
library(lubridate)
library(feasts)
```

## 3.1

```{r}
tsibbledata::global_economy %>%
  filter(Country == 'Sweden') %>%
  autoplot(GDP)
```


```{r}
fit <- global_economy %>%
  model(trend_model = TSLM(GDP ~ trend()))

fit
```

```{r}
fit %>%
  forecast(h = "3 years")
```

```{r}
fit %>%
  forecast(h = "3 years") %>%
  filter(Country == 'Sweden') %>%
  autoplot(global_economy)
```


## 3.2

```{r}
bricks <- aus_production %>%
  filter_index(1970 ~ 2004)

bricks
```

Total Average model

```{r}
bricks %>%
  model(MEAN(Bricks)) %>%
  forecast(h = "5 years") %>%
  autoplot(bricks)
```

Naive: last value

```{r}
bricks %>%
  model(NAIVE(Bricks)) %>%
  forecast(h = "5 years") %>%
  autoplot(bricks)
```

```{r}
bricks %>%
  model(SNAIVE(Bricks ~ lag("year"))) %>%
  forecast(h = "5 years") %>%
  autoplot(bricks)
```

Drift

```{r}
bricks %>%
  model(RW(Bricks ~ drift())) %>%
  forecast(h = "5 years") %>%
  autoplot(bricks)
```

```{r}
# Set training data from 1992 to 2006
train <- aus_production %>% filter_index("1992 Q1" ~ "2006 Q4")

# Fit the models
beer_fit <- train %>%
  model(
    Mean = MEAN(Beer),
    `Naïve` = NAIVE(Beer),
    `Seasonal naïve` = SNAIVE(Beer)
  )

# Generate forecasts for 14 quarters
beer_fc <- beer_fit %>% forecast(h = 14)

# Plot forecasts against actual values
beer_fc %>%
  autoplot(train, level = NULL) +
    autolayer(filter_index(aus_production, "2007 Q1" ~ .), color = "black") +
    ggtitle("Forecasts for quarterly beer production") +
    xlab("Year") + ylab("Megalitres") +
    guides(colour=guide_legend(title="Forecast"))
```

```{r}
# Re-index based on trading days
google_2015 <- gafa_stock %>%
  filter(Symbol == "GOOG") %>%
  mutate(day = row_number()) %>%
  update_tsibble(index = day, regular = TRUE)
```


```{r}
# Filter the year of interest
google_2015 <- google_stock %>% filter(year(Date) == 2015)

# Fit the models
google_fit <- google_2015 %>%
  model(
    Mean = MEAN(Close),
    `Naïve` = NAIVE(Close),
    Drift = NAIVE(Close ~ drift())
  )

# Produce forecasts for the 19 trading days in January 2015
google_fc <- google_fit %>% forecast(h = 19)

# A better way using a tsibble to determine the forecast horizons
google_jan_2016 <- google_stock %>%
  filter(yearmonth(Date) == yearmonth("2016 Jan"))

google_fc <- google_fit %>% forecast(google_jan_2016)

# Plot the forecasts
google_fc %>%
  autoplot(google_2015, level = NULL) +
    autolayer(google_jan_2016, Close, color='black') +
    ggtitle("Google stock (daily ending 31 Dec 2015)") +
    xlab("Day") + ylab("Closing Price (US$)") +
    guides(colour=guide_legend(title="Forecast"))
```

## 3.3

```{r}
global_economy %>%
  filter(Country == "Australia") %>%
  autoplot(GDP / Population)
```

```{r}
print_retail <- aus_retail %>%
  filter(Industry == "Newspaper and book retailing") %>%
  group_by(Industry) %>%
  index_by(Year = year(Month)) %>%
  summarise(Turnover = sum(Turnover))
```

```{r}
aus_economy <- global_economy %>%
  filter(Code == "AUS")

adj_turnover <- print_retail %>%
  left_join(aus_economy, by = "Year") %>%
  mutate(Adjusted_turnover = Turnover / CPI)

adj_turnover %>%
  tidyr::gather("Type", "Turnover", Turnover, Adjusted_turnover, factor_key = TRUE) %>%
  ggplot(aes(x = Year, y = Turnover)) +
    geom_line() +
    facet_grid(vars(Type), scales = "free_y") +
    xlab("Years") + ylab(NULL) +
    ggtitle("Turnover for the Australian print media industry")
```

## 3.4

For example: subtracting out seasonal trends (fitting a model on the months)
and seeing an increase in the residuals.

```{r}
augment(beer_fit)
```

## 3.5

```{r}
google_2015 %>%
  model(NAIVE(Close)) %>%
  forecast(h = 10) %>%
  hilo()
```

```{r}
google_2015 %>%
  model(NAIVE(Close)) %>%
  forecast(h = 10) %>%
  autoplot(google_2015)
```

## Exercises

For the following series, find an appropriate Box-Cox transformation in order to stabilise the variance.

United States GDP from global_economy
Slaughter of Victorian “Bulls, bullocks and steers” in aus_livestock
Gas production from aus_production

```{r}
global_economy %>%
  filter(Country == 'United States') %>% 
  autoplot(GDP)
```

```{r}
global_economy %>%
  filter(Country == 'United States') %>%
  features(GDP, features = guerrero) %>%
  pull(lambda_guerrero)

global_economy %>%
  filter(Country == 'United States') %>% 
  autoplot(box_cox(GDP, 2))
```

```{r}
aus_livestock %>%
  filter(Animal == 'Bulls, bullocks and steers', State == 'Victoria') %>%
  autoplot()
```

```{r}
aus_livestock %>%
  filter(Animal == 'Bulls, bullocks and steers', State == 'Victoria') %>%
  features(features = guerrero) %>%
  pull(lambda_guerrero)

aus_livestock %>%
  filter(Animal == 'Bulls, bullocks and steers', State == 'Victoria') %>% 
  autoplot(box_cox(Count, 2))
```

```{r}
aus_production %>%
  autoplot(box_cox(Gas, 0.2))
```

### e3

```{r}
expsmooth::cangas %>%
  as_tsibble() %>%
  autoplot()
```

```{r}
expsmooth::cangas %>%
  as_tsibble() %>%
  features(value, features = guerrero) %>%
  pull(lambda_guerrero)

expsmooth::cangas %>%
  as_tsibble() %>%
  autoplot(box_cox(value, .4))
```

```{r}
expsmooth::cangas %>%
  as_tsibble() %>%
  gg_season()
```

The season patterns change (more frequent turns as the years increase).

### e4

```{r}
aus_production %>%
  autoplot(Tobacco)
```

```{r}
ansett %>%
  filter(Class == 'Economy') %>%
  as_tibble() %>%
  distinct(Airports)
```


```{r}
ansett %>%
  filter(Class == 'Economy', Airports == 'MEL-SYD') %>%
  autoplot()
```

Demand / population (don't have data)
Demand / temperature

```{r}
vic_elec
```

```{r}
vic_elec %>%
  autoplot(Demand) + 
  scale_y_log10()
```


```{r}
vic_elec %>%
  autoplot(Demand / Temperature) +
  scale_y_log10()
```

### e5

```{r}
global_economy %>%
  filter(Country == 'Australia') %>%
  model(
    NAIVE(Population),
    SNAIVE(Population),
    RW(Population ~ drift())
  ) %>%
  forecast(h = "5 years") %>%
  autoplot(global_economy %>%
  filter(Country == 'Australia'))
```

```{r}
## SNAIVE
aus_production %>%
  autoplot(Bricks)
```

```{r}
brick_fit <- aus_production %>%
  filter(!is.na(Bricks)) %>%
  model(
    Mean = MEAN(Bricks),
    `Naïve` = NAIVE(Bricks),
    `Seasonal naïve` = SNAIVE(Bricks)
  )

# Generate forecasts for 14 quarters
brick_fc <- brick_fit %>% forecast(h=14)

# Plot forecasts against actual values
brick_fc %>%
  autoplot(aus_production %>% filter(!is.na(Bricks))) 
```

```{r}
aus_livestock %>% as_tibble() %>% distinct(Animal)
```

```{r}
aus_livestock %>% 
  filter(Animal == 'Lambs', State == 'New South Wales') %>%
  autoplot()
```

### e6

```{r}
gafa_stock %>% as_tibble() %>% distinct(Symbol)
```

```{r}
filter(gafa_stock, Symbol == 'FB') %>%
  autoplot()
```

```{r}
fb_stock <- filter(gafa_stock, Symbol == 'FB') %>%
  mutate(day = row_number()) %>%
  update_tsibble(index = day, regular = TRUE)
  
fb_stock %>%
  model(
    Mean = MEAN(Open),
    `Naïve` = NAIVE(Open),
    drift = RW(Open ~ drift())
  ) %>%
  forecast(h = 100) %>%
  autoplot(fb_stock)
```



### e7

```{r}
fit <- aus_livestock %>%
  filter(State == 'Victoria') %>%
  model(SNAIVE(Count ~ lag("year"))) 

fit %>%
  forecast(h = "3 years") %>%
  autoplot(aus_livestock %>%
  filter(State == 'Victoria'))
```

Calves seem the best fit. Seems like a reasonable benchmark, since they seem
pretty stable (low drift).


