---
title: "fable-chapter-9"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{fable-chapter-9}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup}
library(tsibble)
library(dplyr)
library(fable)
library(tsibbledata)
library(ggplot2)
library(lubridate)
library(feasts)
library(tidyr)
```

## 9.1

Time series with trends or seasonality are *not* stationary.
- white noise *is* stationary.

Cyclic behavior is stationary
- cycles are not of fixed length 

Expectations
- plot is horizational
- constant variance

```{r}
google_2015 <- gafa_stock %>%
  filter(Symbol == "GOOG") %>%
  mutate(day = row_number()) %>%
  update_tsibble(index = day, regular = TRUE) %>%
  filter(year(Date) == 2015)
```

```{r}
google_2015 %>%
  autoplot(Close)
```


```{r}
google_2015_diff <- google_2015 %>%
  mutate(diff_close = difference(Close))

google_2015_diff %>%
  autoplot(diff_close)
```

```{r}
google_2015_diff %>%
  ACF(diff_close) %>%
  autoplot()
```

```{r}
google_2015_diff %>%
  features(diff_close, ljung_box, lag = 10)
```

```{r}
PBS
```

Sequential seasonal -> first diffence to get stationary series.

```{r}
PBS %>%
  filter(ATC2 == "H02") %>%
  summarise(Cost = sum(Cost)/1e6) %>%
  transmute(
    `Sales ($million)` = Cost,
    `Log sales` = log(Cost),
    `Annual change in log sales` = difference(log(Cost), 12),
    `Doubly differenced log sales` = difference(difference(log(Cost), 12), 1)
  ) %>%
  gather("Type", "Sales", !!!syms(measured_vars(.)), factor_key = TRUE) %>%
  ggplot(aes(x = Month, y = Sales)) +
  geom_line() +
  facet_grid(vars(Type), scales = "free_y") +
  labs(title = "Corticosteroid drug sales", x = "Year", y = NULL)
```

First differences 
- the change between one observation and the next.

Seasonal differences 
- the change between one year to the next

```{r}
google_2015_diff %>%
  features(Close, unitroot_kpss)
```

```{r}
google_2015_diff %>%
  features(diff_close, unitroot_kpss)
```

Determine number of differences to take:

```{r}
google_2015_diff %>%
  features(Close, unitroot_ndiffs)
```

Find number of seasonal diffs

```{r}
aus_total_retail <- aus_retail %>%
  summarise(Turnover = sum(Turnover))


aus_total_retail %>%
  mutate(log_turnover = log(Turnover)) %>%
  features(log_turnover, unitroot_nsdiffs)
```

Then see if first diffs are needed:

```{r}
aus_total_retail %>%
  mutate(log_turnover = difference(log(Turnover), 12)) %>%
  features(log_turnover, unitroot_ndiffs)
```

## 9.3

multiple regression model
- forecast the variable of interest using a linear combination of predictors

autoregression model
- forecast the variable of interest using a linear combination of past values of the variable.

We normally restrict autoregressive models to stationary data

## 9.4

Moving average smoothing vs. moving average models

* moving average model is used for forecasting future values
  * only uses previous forecast errors

* moving average smoothing is used for estimating the trend-cycle of past values
  * looks ahead to next value


## 9.5 

ARIMA 
* acronym for AutoRegressive Integrated Moving Average 
* combinaton of differencing with autoregression and a moving average model

ARIMA(p, d, q)
- p:	order of the autoregressive part;
- d: degree of first differencing involved;
- q: order of the moving average part.

```{r}
us_change <- readr::read_csv(
  "https://otexts.com/fpp3/extrafiles/us_change.csv"
) %>%
  mutate(Time = yearquarter(Time)) %>%
  as_tsibble(index = Time)

us_change
```

No apparent seasonal patterns, and no trend. Seems to be stationary.

```{r}
us_change %>%
  autoplot(Consumption)
```

```{r}
us_change %>%
  features(Consumption, unitroot_kpss)
```

```{r}
us_change %>%
  features(Consumption, unitroot_ndiffs)
```

`PDQ` estimates seasonal parameters. We are ignoring them for this model.

The default for the non-seasonal components are
```
pdq(p = 0:5, d = 0:2, q = 0:5,
    p_init = 2, q_init = 2)
```
One will be automatically chosen to minize `ic` 
(information criteria, AIC, AICc, BIC, etc).

```{r}
fit <- us_change %>%
  model(ARIMA(Consumption ~ PDQ(0,0,0)))

report(fit)
```

constant `c != 0` and `d = 0`, so forecast will trend to mean.

```{r}
fit %>% 
  forecast(h=10) %>% 
  autoplot(slice(us_change, (n()-80):n()))
```

```{r}
us_change %>%
  model(ARIMA(Consumption ~ PDQ(0,0,0) + 0)) %>%
  report()
```

`c = 0` and `d = 0` so forecast will trend to 0.

```{r}
us_change %>%
  model(ARIMA(Consumption ~ PDQ(0,0,0) + 0)) %>%
  forecast(h=10) %>% 
  autoplot(slice(us_change, (n()-80):n()))
```

Finding `p` manually from plots:

```{r}
us_change %>% ACF(Consumption) %>% autoplot()
```

```{r}
us_change %>% PACF(Consumption) %>% autoplot()
```

Look for:

- there is a significant spike at lag p in the PACF, but none beyond lag p

This implies `p = 3`.

Indeed,

```{r}
report(fit)
```

```{r}
fit2 <- us_change %>%
  model(ARIMA(Consumption ~ pdq(3,0,0) + PDQ(0,0,0)))

report(fit2)
```

We get better `ic`s with this model.


We can also force a comprehensive search of `pdq` values:

```{r}
fit3 <- us_change %>%
  model(ARIMA(Consumption ~ PDQ(0,0,0),
              stepwise = FALSE, approximation = FALSE))
report(fit3)
```

And it agrees with the `PACF` plot.

## 9.7

We want to find an ARIMA model on the seasonally-adjusted series.


```{r}
elec_equip <- as_tsibble(fpp2::elecequip)
elec_dcmp <- elec_equip %>%
  model(STL(value ~ season(window="periodic"))) %>%
  components() %>%
  select(-.model) %>%
  as_tsibble()

elec_dcmp %>%
  autoplot(season_adjust)
```

Non stationary. Let's check first difference:

```{r}
elec_dcmp %>%
  gg_tsdisplay(difference(season_adjust), plot_type='partial')
```

PACF -> p = 3.

The difference looks stationary, -> d = 1.

So ARIMA(3, 1, 0) is an option.


```{r}
selection <- elec_dcmp %>%
  model(
    a_310 = ARIMA(season_adjust ~ pdq(3,1,0) + PDQ(0,0,0)),
    a_410 = ARIMA(season_adjust ~ pdq(4,1,0) + PDQ(0,0,0)),
    a_210 = ARIMA(season_adjust ~ pdq(2,1,0) + PDQ(0,0,0)),
    a_311 = ARIMA(season_adjust ~ pdq(3,1,1) + PDQ(0,0,0))
  )

selection %>%
  glance()
```

```{r}
fit <- selection %>%
  select(a_311)
```

```{r}
fit %>%
  gg_tsresiduals()
```

```{r}
augment(fit) %>%
  features(.resid, ljung_box, lag = 24, dof = 4)
```

```{r}
report(fit)
```

`c = 0` and `d = 1`. Forecasts will go to a non-zero constant
(not the mean of the series).

```{r}
fit %>% 
  forecast() %>% 
  autoplot(elec_dcmp) +
  geom_hline(yintercept = mean(elec_dcmp$season_adjust), color = 'red')
```

```{r}
gg_arma(fit)
```

## 9.9

```{r}
eu_retail <- as_tsibble(fpp2::euretail)
eu_retail %>% autoplot(value) + ylab("Retail index") + xlab("Year")
```

```{r}
eu_retail %>%
  gg_season()
```


```{r}
eu_retail %>% gg_tsdisplay(difference(value, 4))
```

Need to do a first difference:

```{r}
eu_retail %>% gg_tsdisplay(value %>% difference(4) %>% difference())
```

Spike at acf 4 -> `Q = 1` (seasonl period 4.)

Spike at acf 1 -> `q = 1`.

-> 

ARIMA(0,1,1)(0,1,1)

One non-seasonal difference and spike at 1,
One seasonal difference and spike at 1 (4 first quarter).

```{r}
eu_retail %>% 
  gg_tsdisplay(value %>% difference(4) %>% difference(), plot_type = 'partial')
```

PACF plot also implies `ARIMA(1,1,0)(1,1,0)`.

We'll try both and compare.

```{r}
fit <- eu_retail %>%
  model(
    a1 = ARIMA(value ~ pdq(0,1,1) + PDQ(0,1,1)),
    a2 = ARIMA(value ~ pdq(0,1,2) + PDQ(0,1,1)),
    a3 = ARIMA(value ~ pdq(0,1,3) + PDQ(0,1,1)),
    a4 = ARIMA(value ~ pdq(1,1,0) + PDQ(1,1,0))
  )

fit %>%
  glance
```

```{r}
fit %>%
  select(a1) %>%
  gg_tsresiduals()
```

```{r}
fit %>%
  select(a3) %>%
  gg_tsresiduals()
```

```{r}
fit %>%
  select(a3, a4) %>%
  forecast(h = 12) %>%
  autoplot(eu_retail, level = NULL)
```

```{r}
eu_retail %>%
  model(ARIMA(value)) %>%
  report
```

```{r}
h02 <- tsibbledata::PBS %>%
  filter(ATC2 == "H02") %>%
  summarise(Cost = sum(Cost)/1e6)

h02 %>%
  gg_tsdisplay()
```

Use `log(Cost)` to make constant variance.

```{r}
h02 %>%
  mutate(log(Cost)) %>%
  gather() %>%
  ggplot(aes(x = Month, y = value)) +
  geom_line() +
  facet_grid(key ~ ., scales = "free_y") +
  xlab("Year") + ylab("") +
  ggtitle("Cortecosteroid drug scripts (H02)")
```

```{r}
h02 %>% 
  gg_tsdisplay(difference(log(Cost), 12), plot_type = 'partial', lag_max = 24)
```

PACF
- non-seasonal 3 spikes -> p = 3
- seasonal at 12 and 24 -> P = 2

```{r}
h02 %>%
  features(difference(log(Cost), 12), unitroot_kpss)
```

```{r}
h02 %>%
  features(difference(log(Cost), 12), unitroot_ndiffs)
```


```{r}
h02 %>% 
  gg_tsdisplay(
    difference(log(Cost), 12) %>% difference, 
    plot_type = 'partial',
    lag_max = 24
  )
```

In `d=2`:
ACF -> 
- non-seasonal q = 1
- season q = 2

```{r}
fit <- h02 %>%
  model(
    std = ARIMA(log(Cost) ~ 0 + pdq(3,0,1) + PDQ(0,1,2))
  )

report(fit)
```

```{r}
fit %>%
  gg_tsresiduals(lag_max = 36)
```

Fails
```{r}
augment(fit) %>%
  features(.resid, ljung_box, lag = 36, dof = 6)
```


Can't compare AICc with multiple diffs

```{r}
h02 %>%
  model(ARIMA(log(Cost))) %>%
  report
```

```{r}
h02 %>%
  model(ARIMA(log(Cost) ~ 0 + pdq(3,0,1) + PDQ(0,1,2))) %>%
  forecast() %>%
  autoplot(h02) +
    ylab("H02 sales (million scripts)") + xlab("Year")
```

## 9.10


```{r}
aus_economy <- global_economy %>% filter(Code == "AUS") %>%
  mutate(Population = Population/1e6)

aus_economy %>%
  gg_tsdisplay(Population)
```

```{r}
aus_economy %>%
  slice(-n()) %>%
  stretch_tsibble(.init = 10) %>%
  model(
    ETS(Population),
    ARIMA(Population)
  ) %>%
  forecast(h = 1) %>%
  accuracy(aus_economy)
```

```{r}
aus_models <- aus_economy %>%
  slice(-n()) %>%
  stretch_tsibble(.init = 10) %>%
  model(
    ets = ETS(Population),
    arima = ARIMA(Population)
  )
```

```{r}
aus_models %>%
  slice(1) %>%
  select(arima) %>%
  report
```




```{r}
aus_economy %>%
  model(ETS(Population)) %>%
  forecast(h = "5 years") %>%
  autoplot(aus_economy)
```

```{r}
# Consider the cement data beginning in 1988
cement <- aus_production %>%
  filter(year(Quarter) >= 1988)

# Use 20 years of the data as the training set
train <- cement %>%
  filter(year(Quarter) <= 2007)
```

```{r}
fit_arima <- train %>% model(ARIMA(Cement))
report(fit_arima)
```

```{r}
gg_tsresiduals(fit_arima, lag_max = 16)
```

```{r}
augment(fit_arima) %>%
  features(.resid, ljung_box, lag = 16, dof = 5)
```

```{r}
fit_ets <- train %>% model(ETS(Cement))
report(fit_ets)
```

```{r}
augment(fit_ets) %>%
  features(.resid, ljung_box, lag = 16, dof = 6)
```

```{r}
# Generate forecasts and compare accuracy over the test set
bind_rows(
  fit_arima %>% 
    accuracy(),
  fit_ets %>% 
    accuracy(),
  fit_arima %>% 
    forecast(h = "2 years 6 months") %>%
    accuracy(cement),
  fit_ets %>% 
    forecast(h = "2 years 6 months") %>%
    accuracy(cement)
)
```

```{r}
# Generate forecasts from an ETS model
fit_ets %>% forecast(h="3 years") %>% autoplot(cement)
```

```{r}
# Generate forecasts from an ETS model
fit_arima %>% forecast(h="3 years") %>% autoplot(cement)
```

