---
title: "fable-chapter-6"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{fable-chapter-6}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r}
library(tsibble)
library(dplyr)
library(fable)
library(tsibbledata)
library(ggplot2)
library(lubridate)
library(feasts)
```

## 6.1

```{r}
fpp3::us_change
```


```{r}
fpp3::us_change %>%
  ggplot(aes(x = Quarter)) +
  geom_line(aes(y = Consumption, colour = "Consumption")) +
  geom_line(aes(y = Income, colour = "Income")) +
  ylab("% change") + xlab("Year") +
  guides(colour=guide_legend(title="series"))
```

```{r}
fpp3::us_change %>%
  ggplot(aes(Income, Consumption)) +
  geom_point() +
  geom_smooth(method = 'lm')
```

```{r}
fpp3::us_change %>%
  model(TSLM(Consumption ~ Income)) %>%
  report()
```

## 6.2

```{r}
fit.consMR <- fpp3::us_change %>%
  model(
    tslm = TSLM(Consumption ~ Income + Production + Unemployment + Savings)
  )

report(fit.consMR)
```

```{r}
augment(fit.consMR)
```

```{r}
augment(fit.consMR) %>%
  ggplot(aes(x = Quarter)) +
  geom_line(aes(y = Consumption, colour = "Data")) +
  geom_line(aes(y = .fitted, colour = "Fitted")) +
  xlab("Year") + ylab(NULL) +
  ggtitle("Percent change in US consumption expenditure") +
  guides(colour=guide_legend(title=NULL))
```


```{r}
augment(fit.consMR) %>%
  ggplot(aes(x=Consumption, y=.fitted)) +
  geom_point() +
  ylab("Fitted (predicted values)") +
  xlab("Data (actual values)") +
  ggtitle("Percent change in US consumption expenditure") +
  geom_abline(intercept=0, slope=1)
```

## 6.3

```{r}
fit.consMR %>%
  gg_tsresiduals()
```


```{r}
augment(fit.consMR) %>%
  ggplot(aes(x=.fitted, y=.resid)) +
  geom_point() +
  labs(x = "Fitted", y = "Residuals")
```

## 6.4

```{r}
aus_production
```

```{r}
recent_production <- aus_production %>%
  filter(year(Quarter) >= 1992)

recent_production %>%
  autoplot(Beer) +
  labs(x = "Year", y = "Megalitres")
```

```{r}
recent_production %>%
  gg_tsdisplay(Beer)
```

```{r}
fit_beer <- recent_production %>%
  model(TSLM(Beer ~ trend() + season()))

report(fit_beer)
```

```{r}
fit_beer %>%
  forecast() %>%
  autoplot(recent_production)
```

```{r}
augment(fit_beer)
```

```{r}
augment(fit_beer) %>%
  ggplot(aes(Quarter)) +
  geom_line(aes(y = Beer, color = "actual")) +
  geom_line(aes(y = .fitted, color = 'fitted'))
```

```{r}
fit_beer %>%
  gg_tsresiduals()
```

```{r}
augment(fit_beer) %>%
  ggplot(aes(x = Beer, y = .fitted,
             colour = factor(quarter(Quarter)))) +
  geom_point() +
  ylab("Fitted") + xlab("Actual values") +
  ggtitle("Quarterly beer production") +
  scale_colour_brewer(palette="Dark2", name="Quarter") +
  geom_abline(intercept=0, slope=1)
```

```{r}
fourier_beer <- recent_production %>%
  model(TSLM(Beer ~ trend() + fourier(K=2)))

report(fourier_beer)
```

## 6.5

```{r}
glance(fit.consMR)
```

```{r}
glance(fit.consMR) %>%
  select(adj_r_squared, CV, AIC, AICc, BIC)
```

## 6.6

No difference between ex-ante and ex-post forecasts, 
because the model only depends on trend and seasonality. 
There are no predictors to forecast, like Consumption example.

```{r}
recent_production <- aus_production %>% filter(year(Quarter) >= 1992)

fit_beer <- recent_production %>%
  model(TSLM(Beer ~ trend() + season()))


fc_beer <- forecast(fit_beer)
fc_beer %>%
  autoplot(recent_production) +
  ggtitle("Forecasts of beer production using regression") +
  xlab("Year") + ylab("megalitres")
```

We can do a ex-ante forecast by adding conversation predictions are variables
of interest to see how our model behaves.

For example, let say we are working with
```{r}
fpp3::us_change
```

And we want to make forecasts based on these growth patterns for the
predictors:

```{r}
up_future <- new_data(fpp3::us_change, 4) %>%
  mutate(Income = 1, Savings = 0.5, Unemployment = 0)

up_future
```

```{r}
down_future <- new_data(fpp3::us_change, 4) %>%
  mutate(Income = -1, Savings = -0.5, Unemployment = 0)

down_future
```

Build model:

```{r}
fit_consBest <- fpp3::us_change %>%
  model(tslm = TSLM(Consumption ~ Income + Savings + Unemployment))

report(fit_consBest)
```

```{r}
fc_up <- forecast(fit_consBest, new_data = up_future) %>%
  mutate(Scenario = 'Increase') %>%
  as_fable(response=Consumption, key = Scenario)

fc_up
```

```{r}
fc_down <- forecast(fit_consBest, new_data = down_future) %>%
  mutate(Scenario = 'Decrease') %>%
  as_fable(response=Consumption, key = Scenario)

fc_down
```

```{r}
fpp3::us_change %>%
  autoplot(Consumption) +
  autolayer(rbind(fc_up, fc_down)) +
  ylab("% change in US consumption")
```

This is still *NOT* ex-post, because we don't have true values for the predictors.

```{r}
trend_model <- fpp3::us_change %>%
  model(tslm = TSLM(Consumption ~ trend())) 

forecast(trend_model) %>%
  autoplot(fpp3::us_change)
```

## 6.8

Not linear trends, piece-wise linear.

```{r}
boston_men <- fpp3::boston_marathon %>%
  filter(Event == "Men's open division") %>%
  mutate(Minutes = as.numeric(Time)/60)

boston_men
```

```{r}
boston_men %>%
  autoplot(Minutes) +
  geom_smooth(method = 'lm')
```

```{r}
fit_trends <- boston_men %>%
  model(
    linear = TSLM(Minutes ~ trend()),
    exponential = TSLM(log(Minutes) ~ trend()),
    piecewise = TSLM(Minutes ~ trend(knots = c(1940, 1980)))
  )

fit_trends %>%
  select(piecewise) %>%
  report()
```

```{r}
fc_trends <- fit_trends %>% forecast(h=10)
```


```{r}
boston_men %>%
  autoplot(Minutes) +
  geom_line(aes(y=.fitted, colour=.model), data = fitted(fit_trends)) +
  autolayer(fc_trends, alpha = 0.5, level = 95) +
  xlab("Year") +  ylab("Winning times in minutes") +
  ggtitle("Boston Marathon") +
  guides(colour=guide_legend(title="Model"))
```

## 6.10 exercises

### 1

```{r}
jan14_vic_elec <- tsibbledata::vic_elec %>%
  filter(yearmonth(Time) == yearmonth("2014 Jan")) %>%
  index_by(Date = as_date(Time)) %>%
  summarise(Demand = sum(Demand), Temperature = max(Temperature))

jan14_vic_elec
```

#### a

```{r}
jan14_vic_elec %>%
  autoplot()
```

```{r}
fit_elec <- jan14_vic_elec %>%
  model(linear = TSLM(Demand ~ Temperature))

report(fit_elec)
```

```{r}
augment(fit_elec) %>%
  ggplot(aes(Date)) +
  geom_line(aes(y = Demand, color = "True")) +
  geom_line(aes(y = .fitted, color = "Fit"))
```

#### b

```{r}
fit_elec %>%
  gg_tsresiduals()
```

```{r}
temp_15 <- new_data(jan14_vic_elec, 1) %>%
  mutate(Temperature = 15)
temp_15
```

```{r}
temp_35 <- new_data(jan14_vic_elec, 1) %>%
  mutate(Temperature = 35)
```

```{r}
fc_15 <- fit_elec %>%
  forecast(new_data = temp_15) %>%
  mutate(fc = 'low 15') %>%
  as_fable(response = Demand, key = fc)
```

```{r}
fc_35 <- fit_elec %>%
  forecast(new_data = temp_35) %>%
  mutate(fc = 'high 35') %>%
  as_fable(response = Demand, key = fc)
```

15 degrees will be below observated value, 
so the prediced demand will have a larger error.

```{r}
jan14_vic_elec %>%
  as_tibble() %>%
  ggplot(aes(Temperature)) +
  geom_histogram()
```

```{r}
jan14_vic_elec %>%
  autoplot(Demand) +
  autolayer(rbind(fc_15, fc_35))
```

Hilo diff: 202683 - 100585 = 102,098

```{r}
fc_15 %>%
  mutate(interval = hilo(.distribution))
```

Hilo diff: 324070 - 225260 = 98,810

```{r}
fc_35 %>%
  mutate(interval = hilo(.distribution))
```

So the prediction interval for the most likely temperature (35) is small,
but not by that much. 

#### e

```{r}
jan14_vic_elec %>%
  as_tibble() %>%
  ggplot(aes(Temperature, Demand)) +
  geom_point() + 
  geom_smooth(method = 'lm')
```

### 2

```{r}
tsibbledata::olympic_running
```

```{r}
tsibbledata::olympic_running %>%
  as_tibble() %>%
  group_by(Length, Sex) %>%
  summarise(n = n())
```


```{r}
tsibbledata::olympic_running %>%
  filter(Length == '100m') %>%
  autoplot()
```

```{r}
fit_running <- tsibbledata::olympic_running %>%
  filter(Length == '100m') %>%
  model(linear = TSLM(Time ~ trend()))
```

```{r}
fit_running %>%
  forecast() %>%
  autoplot(tsibbledata::olympic_running %>% filter(Length == '100m'))
```

```{r}
augment(fit_running) %>%
  ggplot(aes(Year, .resid, color = Sex)) +
  geom_point()
```

```{r}
fit_events <- tsibbledata::olympic_running %>%
  model(tslm = TSLM(Time ~ trend()))

fit_events
```

```{r}
fit_events %>%
  forecast(h = 1) %>% 
  as_tibble() %>%
  mutate(interval = hilo(.distribution)) %>%
  select(Length, Sex, Time, interval)
```

### 4

```{r}
fancy <- fma::fancy %>%
  as_tsibble()
fancy %>%
  autoplot()
```

#### a

```{r}
fancy %>%
  gg_tsdisplay()
```

```{r}
fit_fancy_trend <- fancy %>%
  model(
    linear = TSLM(value ~ trend()),
    log = TSLM(log(value) ~ trend())
  )

augment(fit_fancy_trend)
```

```{r}
accuracy(fit_fancy)
```


```{r}
augment(fit_fancy_trend) %>%
  ggplot(aes(index)) +
  geom_line(aes(y = value, color = "value")) +
  geom_line(aes(y = .fitted, color = .model))
```

The residuals of the linear trend are not centered around zero.

```{r}
fit_fancy_trend %>%
  select(linear) %>%
  gg_tsresiduals()
```

The log trend is centered around zero.

```{r}
fit_fancy_trend %>%
  select(log) %>%
  gg_tsresiduals()
```

Adding in seasonality, the log fit is much, much better than the linear.

```{r}
fit_fancy_seasonal <- fancy %>%
  model(
    linear = TSLM(value ~ trend() + season()),
    log = TSLM(log(value) ~ trend() + season())
  )

accuracy(fit_fancy_seasonal)
```

```{r}
fit_fancy_seasonal %>%
  select(linear) %>%
  gg_tsresiduals()
```

```{r}
fit_fancy_seasonal %>%
  select(log) %>%
  gg_tsresiduals()
```

This explains the major difference between the log and linear model.

As you can see by the red line plot, the linear forecast expects linear growth,
but that doesn't scale with the increase of sales.

With the `log` fit, the trend or growth is exponential, 
and the fit finds the right exponential growth coefficient. 

This means it is able to match the increase in volume in later years.

```{r}
augment(fit_fancy_seasonal) %>%
  ggplot(aes(index)) +
  geom_line(aes(y = value, color = "value")) +
  geom_line(aes(y = .fitted, color = .model))
```

#### c

```{r}
surf_fancy <- fancy %>%
  mutate(surf = as.factor(case_when(
    month(index) == 3 & year(index) > 1987 ~ "yes", 
    TRUE ~ "no"
  )))

surf_fancy
```

```{r}
fit_fancy_surf <- surf_fancy %>%
  model(
    seasonal = TSLM(log(value) ~ trend() + season()),
    surf = TSLM(log(value) ~ trend() + season() + surf)
  )

accuracy(fit_fancy_surf)
```

```{r}
fit_fancy_surf %>%
  select(surf) %>%
  report()
```

```{r}
augment(fit_fancy_surf) %>%
  ggplot(aes(index)) +
  geom_line(aes(y = value, color = "value")) +
  geom_line(aes(y = .fitted, color = .model))
```

#### d

```{r}
fit_fancy_surf %>%
  select(surf) %>%
  gg_tsresiduals()
```

```{r}
augment(fit_fancy_surf) %>%
  filter(.model == 'surf') %>%
  ggplot(aes(.fitted, .resid)) +
  geom_point()
```

```{r}
augment(fit_fancy_surf) %>%
  mutate(month = month(index, label = TRUE)) %>%
  ggplot(aes(month, .resid, group = month)) +
  geom_boxplot()
```

Therefore, the residuals *are* distinguishable from white noise,
meaning there is signal in the residuals our model is not accounting for.

```{r}
augment(fit_fancy_surf) %>%
  features(.resid, ljung_box)
```


```{r}
augment(fit_fancy_surf) %>%
  features(.resid, ljung_box, lag = 10, dof = 3)
```

```{r}
fit_fancy_surf %>%
  select(seasonal) %>%
  forecast(h = 3 * 12) %>%
  autoplot(fancy)
```

### 5

```{r}
us_gas <- fpp3::us_gasoline %>%
  filter_index(~ "2005")

us_gas
```

```{r}
us_gas %>%
  gg_tsdisplay()
```


```{r}
fit_gas <- us_gas %>%
  model(
    f2 = TSLM(Barrels ~ trend() + fourier(K = 2)),
    f3 = TSLM(Barrels ~ trend() + fourier(K = 3)),
    f5 = TSLM(Barrels ~ trend() + fourier(K = 5)),
    f9 = TSLM(Barrels ~ trend() + fourier(K = 9)),
    f12 = TSLM(Barrels ~ trend() + fourier(K = 12))
  )

accuracy(fit_gas)
```

```{r}
augment(fit_gas) %>%
  ggplot(aes(Barrels, .fitted)) +
  geom_point() +
  facet_wrap(~.model)
```

```{r}
fit_gas %>%
  select(f12) %>%
  gg_tsresiduals()
```


#### d

```{r}
fit_gas %>%
  select(f2, f12) %>%
  forecast(h = 52) %>%
  autoplot(
    fpp3::us_gasoline %>%
      filter_index("2005" ~ "2006")
  )
```

The more fourier terms, the more wiggly the curve gets.

### 6

```{r}
huron <- fma::huron %>%
  as_tsibble()

huron
```

```{r}
gg_tsdisplay(huron)
```

```{r}
fit_huron <- huron %>%
  model(
    linear = TSLM(value ~ trend()),
    knot = TSLM(value ~ trend(knots = c(1915)))
  )

accuracy(fit_huron)
```

```{r}
augment(fit_huron) %>%
  ggplot(aes(index)) +
  geom_line(aes(y = value, color = "value")) +
  geom_line(aes(y = .fitted, color = .model))
```

```{r}
fit_huron %>%
  forecast(h = 8) %>%
  autoplot(huron)
```

