---
title: "fable-book"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{fable-book}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

## Fable Book


### 2.1

```{r}
library(tsibble)
library(dplyr)

y <- tsibble(Year = 2015:2019, Observation = c(123,39,78,52,110), index = Year)
y
```

Really this is multiple time-series of events and sex:

```{r}
tsibbledata::olympic_running
```

```{r}
tsibbledata::olympic_running %>%
  dplyr::as_tibble() %>%
  dplyr::count(Length, Sex)
```

If their are duplicated rows, `tsibble` will through an error
because it is not a valid time-series.

```{r}
tsibbledata::PBS %>%
  select(Month, Concession, Type, Cost)
```

```{r}
tsibbledata::PBS %>%
  filter(ATC2=="A10") %>%
  select(Month, Concession, Type, Cost)
```

`tsibble` automatically selects month, the index variable.

Also automatically grouped by index:

```{r}
tsibbledata::PBS %>%
  filter(ATC2=="A10") %>%
  select(Month, Concession, Type, Cost) %>%
  summarise(TotalC = sum(Cost))
```

```{r}
a10 <- tsibbledata::PBS %>%
  filter(ATC2=="A10") %>%
  select(Month, Concession, Type, Cost) %>%
  summarise(TotalC = sum(Cost)) %>%
  mutate(Cost = TotalC/1e6)
a10
```


## 2.2

```{r}
tsibbledata::ansett
```

Complicated vs. simple time series

```{r}
melsyd_economy <- tsibbledata::ansett %>%
  filter(Airports == "MEL-SYD", Class=="Economy")

library(ggplot2)

melsyd_economy %>%
  autoplot(Passengers) +
    labs(title = "Ansett economy class passengers", subtitle = "Melbourne-Sydney") +
    xlab("Year")
```

```{r}
a10 %>% autoplot(Cost) +
  ggtitle("Antidiabetic drug sales") +
  ylab("$ million") + xlab("Year")
```

## 2.2

Time series vocab:

### Trend

long-term increase of decrease of data

### Seasonal

changes during a fixed and know period

### Cyclic

changes not at a fixed frequency


## 2.4

Season plots: data plotted in the season that they occured.
> see for example, when Hadley plotted against the year
to look at the decreasing year-end housing permits

```{r}
a10 %>% feasts::gg_season(Cost, labels = "both") +
  ylab("$ million") +
  ggtitle("Seasonal plot: antidiabetic drug sales")
```

These plots show the underlying season pattern more clearly.

```{r}
library(feasts)

a10 %>%
  gg_subseries(Cost) +
    ylab("$ million") +
    xlab("Year") +
    ggtitle("Seasonal subseries plot: antidiabetic drug sales")
```

Shows diffferent seasonality patterns within months.

Also shows changes in seasonality over time (growth through all months over years).


## 2.6

Relationships **between** time series
> see for example, Hadley plotting between cities

Also, plotting series against each other:

```{r}
library(lubridate)

tsibbledata::vic_elec %>%
  filter(year(Time) == 2014) %>%
  autoplot(Demand) +
    xlab("Year: 2014") + ylab(NULL) +
    ggtitle("Half-hourly electricity demand: Victoria, Australia")
```

```{r}
tsibbledata::vic_elec %>%
  filter(year(Time) == 2014) %>%
  autoplot(Temperature) +
    xlab("Year: 2014") + ylab(NULL) +
    ggtitle("Half-hourly electricity demand: Victoria, Australia")
```

We roughly see when demand is high when temperature is low in July.

```{r}
tsibbledata::vic_elec %>%
  filter(year(Time) == 2014) %>%
  ggplot(aes(x = Temperature, y = Demand)) +
    geom_point() +
    ylab("Demand (GW)") + xlab("Temperature (Celsius)")
```

## 2.7

Lag plots

```{r}
recent_production <- tsibbledata::aus_production %>%
  filter(year(Quarter) >= 1992)
recent_production %>% gg_lag(Beer, geom="point")
```

Positive correlation on 4 and 8 indicate seasonality.

## 2.8 

Autocorrelation

These correspond with the correlation we see in Section 2.7 with lag plots.

```{r}
recent_production %>% ACF(Beer, lag_max = 9)
```

We see positive correlation at 4 and 8 and negative correlation at 2 and 6,
and little correlation elsewhere.

```{r}
recent_production %>% ACF(Beer) %>% autoplot()
```

```{r}
a10 %>% ACF(Cost, lag_max = 48) %>% autoplot()
```

## Exercises

Use the help function to explore what the series gafa_stock, PBS, vic_elec and pelt represent.

Use autoplot() to plot some of the series in these data sets.
What is the time interval of each series?
Use filter() to find what days corresponded to the peak closing price for each of the four stocks in gafa_stock.

```{r}
tsibbledata::gafa_stock
```

```{r}
tsibbledata::gafa_stock %>%
  as_tibble() %>%
  distinct(Symbol)
```

```{r}
tsibbledata::gafa_stock %>%
  autoplot()
```

```{r}
tsibbledata::gafa_stock %>%
  as_tibble() %>%
  group_by(Symbol) %>%
  filter(Close == max(Close, na.rm = TRUE)) %>%
  select(Symbol, Date, Close)
```

```{r}
tsibbledata::pelt %>%
  autoplot()
```

### e3

Create time plots of the following time series: fma::bicoal, fma::chicken, fma::dole, USAccDeaths, fma::writing, fma::fancy.

Use ? (or help()) to find out about the data in each series.
You will need to convert each series to a tsibble using as_tsibble().

--

These are

```{r}
fma::bicoal %>%
  class
```

```{r}
?fma::bicoal

fma::bicoal %>%
  as_tsibble() %>%
  autoplot()
```

```{r}
?fma::chicken

fma::chicken %>%
  as_tsibble() %>%
  autoplot()
```

```{r}
?fma::dole

fma::dole %>%
  as_tsibble() %>% 
  autoplot()

fma::dole %>%
  as_tsibble() %>% 
  gg_season() + 
  scale_y_log10()
```

```{r}
?fma::writing

fma::writing %>%
  as_tsibble() %>%
  autoplot()

fma::writing %>%
  as_tsibble() %>%
  gg_season()

fma::writing %>%
  as_tsibble() %>%
  gg_subseries()
```

```{r}
?fma::fancy

fma::fancy %>%
  as_tsibble() %>%
  autoplot()

fma::fancy %>%
  as_tsibble() %>%
  gg_season()

fma::fancy %>%
  as_tsibble() %>%
  gg_subseries()
```

```{r}
a10 %>%
  gg_subseries()
```


### e5

Use the following graphics functions: autoplot(), gg_season(), gg_subseries(), gg_lag(), ACF() and explore features from the following time series: fma::hsales, USAccDeaths, bricks from aus_production, fpp2::sunspotarea, fpp2::gasoline.

Can you spot any seasonality, cyclicity and trend?
What do you learn about the series?

--

In the subseries plot, we see the yearly cyclical pattern across months.

```{r}
fma::hsales %>%
  as_tsibble() %>%
  autoplot()

fma::hsales %>%
  as_tsibble() %>%
  gg_season()

fma::hsales %>%
  as_tsibble() %>%
  gg_subseries()

fma::hsales %>%
  as_tsibble() %>%
  gg_lag()

fma::hsales %>%
  as_tsibble() %>%
  ACF()
```

```{r}
fpp2::sunspotarea %>%
  as_tsibble() %>%
  autoplot()
```

There seems to be a cyclic trend, but not on a seasonal period. 
We can study the lag, the correlation with previous values over time.

We see that that there are negative correlations on a 5-year period,
and positive correlations on a 10-year period. 
Moveover, this cycle continues!

```{r}
fpp2::sunspotarea %>%
  as_tsibble() %>%
  gg_lag()

fpp2::sunspotarea %>%
  as_tsibble() %>%
  ACF() %>%
  autoplot()
```


```{r}
fpp2::gasoline %>%
  as_tsibble() %>%
  autoplot()
```

```{r}
fpp2::gasoline %>%
  as_tsibble() %>%
  gg_season() +
  geom_smooth(aes(group = NULL), se = FALSE)
```


```{r}
tsibbledata::aus_production %>%
  autoplot(Bricks)
```

```{r}
tsibbledata::aus_production %>%
  gg_season()
```

```{r}
tsibbledata::aus_production %>%
  gg_lag() 
```

### e7

```{r}
tsibbledata::pelt %>%
  ACF() %>%
  autoplot()
```


### e8

```{r}
tsibbledata::aus_livestock %>%
  as_tibble() %>%
  distinct(State)

tsibbledata::aus_livestock %>%
  as_tibble() %>%
  distinct(Animal)
```

```{r}
vic_pigs <- tsibbledata::aus_livestock %>%
  filter(State == 'Victoria', Animal == "Pigs") %>%
  filter_index("1990" ~ "1995")

vic_pigs %>%
  autoplot()
```

```{r}
vic_pigs %>%
  ACF() %>%
  autoplot()
```

### e9

```{r}
tsibbledata::gafa_stock %>%
  as_tibble() %>%
  distinct(Symbol)
```

```{r}
goog_diff <- tsibbledata::gafa_stock %>%
  filter(Symbol == 'GOOG') %>%
  mutate(diff = difference(Close))
goog_diff %>%
  autoplot(diff)
```

```{r}
goog_diff %>%
  ACF(diff) %>%
  autoplot()
```

